---
title: "starkr Optimizations"
author: "Brian Richards"
date: "November 18, 2015"
output: html_document
---

# Optimizing the code in starkr

Initial library calls
```{r library calls}
library(starkr)
library(dplyr)
library(ggplot2)
library(profvis)
```

First I need to get a benchmark for how my functions run. Particularly, I want 
to look at the speed at which `stark_matrix()` runs. I assume the bottleneck 
is in `radial_matrix_element()`.

```{r state call}
states <- starkr::state_list(15, 15, 1/2, 0, 0)
n <- states[, 1]
l <- states[, 2]
j <- states[, 3]
m_j <- 1/2
```

```{r initial benchmark}
profvis(starkr::stark_matrix(n, l, j, m_j))
```

The output from here looks like the vast majority of the bottleneck is in fact 
in `radial_matrix_element()` with a small portion of the bottleneck going to 
`clebsch_gordan()` and `sphere_mat_element()`. Now to dive in a little deeper 
with a look specifically at `radial_matrix_element()`.

```{r benchmark of rad mat elem}
profvis(starkr::radial_matrix_element(15, 15, 0, 1, 1/2, 1/2))
```

The bottleneck appears to be specifically in a line of code where a data frame 
is produced. I think that this line is actually no longer used in the code and 
is a left over from a previous iteration. A new version of the function called 
`optim_radial_matrix_element()` has been created with that line commented out.
