---
title: "Introduction to starkr"
author: "Brian Richards"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: bibliography.bib
vignette: >
  %\VignetteIndexEntry{Introduction to starkr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The starkr package is designed to calculate the Stark matrix and Stark maps 
for Rubidium-85. The code relies heavily on techniques and algorithms used by 
[@Zimmerman1979] in their paper analyzing the Stark structure of alkali-metal 
atoms. The code described below is broken up in to several sections:

* The underlying structure of the Stark matrix elements.
* Creating the Stark matrix elements and Stark matrix.
* Diagonalizing the Stark matrix to create a Stark map.
* Data tidying to create an easier to view and manipulate Stark map.
* Analysis of Stark states at particular field values.

## Components of the Stark matrix element
All of the functions that calcuate the underlying pieces of the Stark matrix 
element may be used independently of the Stark matrix element calculation.

### Quantum Defect
`quantum_defect()` calculates the quantum defect of an arbitrary (n, l, j) 
state of Rubidium-85. The quantum defect is calculated using a modified version 
of equation (16.19) from [_insert Gallagher Rydberg Atoms reference_] pg. 351. 

$$\delta_{n, l, j} = \delta_{0} + \frac{\delta_{2}}{(n - \delta_{0})^{2}}$$

The parameters $\delta_{0}$ and $\delta_{2}$ are experimentally measured values 
from the following papers [_insert Refereneces_].

### Clebsch-Gordan Coefficients and Wigner-3j Symbols
`clebsch_gordan()` calculates the Clebsch-Gordan coefficient for an arbitrary 
selection of quantum numbers j~1~, j~2~, m~1~, m~2~, j, and m~j~. The conditions 
that must be met for a Clebsch-Gordan coefficient are that $m_{1} + m_{2} = 
m_{j}$.

There are two functions that calculate the Clebsch-Gordan coefficient within 
starkr. The first is `clebsch_gordan()` which uses an analytic form which can 
be found in Edmonds [_insert reference_] or Lindgren and Morrison 
[_insert Reference_]. The second is `clebsch_gordan2()` which uses an analytic 
form found in Cromwell's _Group Theory of Physics_ [_insert Reference_].

`wigner_3j()` is provided for completeness and is not actually used in the 
calculation of the Stark matrix element. It is calculated using the 
Clebsch-Gordan coefficient. 

[Insert Examples]

### Spherical Matrix Element
`sphere_mat_element()` calculates the spherical matrix element
$<l, m|\cos{\theta}|l', m>$ based on the form given by @Zimmerman1979.

### Radial Matrix Element

## Vignette Info

Note the various macros within the `vignette` section of the metadata block above. These are required in order to instruct R how to build the vignette. Note that you should change the `title` field and the `\VignetteIndexEntry` to match the title of your vignette.

## Styles

The `html_vignette` template includes a basic CSS theme. To override this theme you can specify your own CSS in the document metadata as follows:

    output: 
      rmarkdown::html_vignette:
        css: mystyles.css

## Figures

The figure sizes have been customised so that you can easily put two images side-by-side. 

```{r, fig.show='hold'}
plot(1:10)
plot(10:1)
```

You can enable figure captions by `fig_caption: yes` in YAML:

    output:
      rmarkdown::html_vignette:
        fig_caption: yes

Then you can use the chunk option `fig.cap = "Your figure caption."` in **knitr**.

## More Examples

You can write math expressions, e.g. $Y = X\beta + \epsilon$, footnotes^[A footnote here.], and tables, e.g. using `knitr::kable()`.

```{r, echo=FALSE, results='asis'}
knitr::kable(head(mtcars, 10))
```

Also a quote using `>`:

> "He who gives up [code] safety for [code] speed deserves neither."
([via](https://twitter.com/hadleywickham/status/504368538874703872))
